An investigation into using scagnostics to find interesting projections of high-dimensional data

### 2018-10-22 scagnostics.Rmd
Keep all timers, still need to combine to meaningful plot.

timeThis <- function(d, t, idx, pmax){ #d=data matrix, t=interpolated tour path, idx=index function, pmax=max number of projections
    i <- 1
    tic.clearlog()
    tic() #start timer
    for(pMatrix in t){
      if(i>pmax) break
      dProj <- d %*% pMatrix
      sgnst <- idx(dProj[,1],dProj[,2])
      i = i+1
    }
    toc(log=TRUE,quiet=TRUE)
    scTd <- unlist(tic.log(format=FALSE))["toc.elapsed"]-unlist(tic.log(format=FALSE))["tic.elapsed"]
}

@@

  sizeL <- c(10, 100, 500, 1000, 5000, 10000)
  n <- length(sizeL)
  dfTimer = data.frame( n=rep(0, n), scagT=rep(0,n), dcorT=rep(0,n), splineT=rep(0,n), mineT=rep(0,n),
                        mineeT=rep(0,n))
  t <- 1
  for(sampleSize in sizeL){
    set.seed(sampleSize + 11) # new seed for each sample size
    if(sampleSize == 0) sampleSize = 10 # smallest considered sample has 10 points
    thisMatrix <-  sphereData(6, sampleSize) %>% rescale()
    scagT <- timeThis(thisMatrix, grandTour100, scagnostics.default, 100)
    dcorT <- timeThis(thisMatrix, grandTour100, dcor2d, 100)
    splineT <- timeThis(thisMatrix, grandTour100, splines2d, 100)
    mineT <- timeThis(thisMatrix, grandTour100, mine, 100)
    mineeT <- timeThis(thisMatrix, grandTour100, mineE, 100)
    dfTimer[t,] <- c(sampleSize, scagT, dcorT, splineT, mineT,mineeT)
    t <-  t+1
  }
  save(dfTimer, file = "cache/timer.rda")
} else {
  load("cache/timer.rda")
}

timerMelt <- melt(dfTimer, id="n")
ggplot(timerMelt, aes(x = n, y = value, color=variable)) + 
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "Sample Size", y = "Time [s]", color = "Index family")
```

@@

```{r bbhGuided3, results="hide", fig.cap=paste("Final views identified in the dataset considering the seven dimensional parameter space (alpha, theta\\_jn, chi\\_tot, chi\\_p, ra, dec, distance), differing only by randomly selected starting plane. The correponding index values for the six views are: (", toString(format(idxL, digits=2)), ")."), dev = "png", dpi=300}
bbhM3 <- rescale(select(bbhD, alpha, theta_jn, chi_tot, chi_p, ra, dec, distance))
getResWithSeed <- function(s){
  set.seed(s)
  startM <- orthonormalise(matrix(runif(14),ncol = 2))
  tourRes <- save_history(bbhM3, guided_tour(splineIndex()))
}

if(!file.exists("cache/bbhGuidedSpline3.rda")){
  set.seed(2018)
  seedVals <- sample(1:10000, 6)
  bbhRes1 <- getResWithSeed(seedVals[1])
  bbhRes2 <- getResWithSeed(seedVals[2])
  bbhRes3 <- getResWithSeed(seedVals[3])
  bbhRes4 <- getResWithSeed(seedVals[4])
  bbhRes5 <- getResWithSeed(seedVals[5])
  bbhRes6 <- getResWithSeed(seedVals[6])
  save(bbhRes1, bbhRes2, bbhRes3, bbhRes4, bbhRes5, bbhRes6, file = "cache/bbhGuidedSpline3.rda")
} else {
  load("cache/bbhGuidedSpline3.rda")
}
resList <- list(as.list(bbhRes1), as.list(bbhRes2), as.list(bbhRes3),
                as.list(bbhRes4), as.list(bbhRes5), as.list(bbhRes6))
pList <- list()
idxL <- NULL
i <- 1
xmins <- c(-0.5,-0.5,-0.4,-0.9,-1,0.25)
xmaxs <- c(0.75,0.75,0.8,0.4,0.25,1.5)
ymins <- c(-0.8,-0.8,-0.7,-0.8,-1.2,-0.6)
ymaxs <- c(0.7,0.6,0.7,0.7,0.,0.9)
xscales <- c(0.25,0.25,0.24,0.26,0.25,0.25)
yscales <- c(0.3,0.28,0.28,0.3,0.24,0.3)
xcents <- c(0.25,-0.25,-0.1,0.1,-0.75,1.25)
ycents <- c(0.4,-0.4,-0.4,-0.4,-0.75,0.5)
for (r in resList){
  iLast <- length(r)
  fProj <- r[[iLast]]
  dProj <- as.tibble(bbhM3 %*% fProj)
  colnames(fProj) <- c("PP1", "PP2")
  colnames(dProj) <- c("PP1", "PP2")
  fProj <- as_tibble(fProj) %>%
    add_column(label=colnames(bbhM3))
  idxL <- c(idxL, splines2d(dProj$PP1, dProj$PP2))
  pList[[i]] <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)+ 
  geom_path(data=getCircle(xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x, y=y), color="grey") + 
  xlim(xmins[i], xmaxs[i]) + ylim(ymins[i], ymaxs[i]) +
  geom_segment(data=getAxes(fProj, xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x1, xend=x2, y=y1, yend=y2)) +
  geom_text(data=getAxes(fProj, xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x2, y=y2, label=label), size=2.5)
  i <- i+1
}
grid.arrange(pList[[1]], pList[[2]], pList[[3]],pList[[4]], pList[[5]], pList[[6]], ncol=3)
```

>>>>

timeThis <- function(d, t, idx, pmax, n, idxName){ #d=data matrix, t=interpolated tour path, idx=index function, pmax=max number of projections, n=sample size, idxName=str name of idx funciton
    i <- 1
    dfTimer = data.frame( t= numeric(), i=numeric(), n = numeric(), name=character())
    for(pMatrix in t){
      if(i>pmax) break
      tic.clearlog()
      tic() #start timer
      dProj <- d %*% pMatrix
      sgnst <- idx(dProj[,1],dProj[,2])
      toc(log=TRUE,quiet=TRUE)
      scTd <- unlist(tic.log(format=FALSE))["toc.elapsed"]-unlist(tic.log(format=FALSE))["tic.elapsed"]
      dfTimer <- add_row(dfTimer, t=scTd, i=i, n= n, name=idxName)
      i = i+1
    }
    return(dfTimer)
}

@@

  sizeL <- c(10, 100, 500, 1000, 5000, 10000)
  t <- 1
  dfTimer <- NULL
  for(sampleSize in sizeL){
    set.seed(sampleSize + 11) # new seed for each sample size
    if(sampleSize == 0) sampleSize = 10 # smallest considered sample has 10 points
    thisMatrix <-  sphereData(6, sampleSize) %>% rescale()
    scagT <- timeThis(thisMatrix, grandTour100, scagnostics.default, 100, sampleSize, "scagnostics")
    dcorT <- timeThis(thisMatrix, grandTour100, dcor2d, 100, sampleSize, "dcor2D")
    splineT <- timeThis(thisMatrix, grandTour100, splines2d, 100, sampleSize, "splines2D")
    mineT <- timeThis(thisMatrix, grandTour100, mine, 100, sampleSize, "MINE")
    mineeT <- timeThis(thisMatrix, grandTour100, mineE, 100, sampleSize, "MINE E")
    dfTimerC <- list(scagT, dcorT, splineT, mineT,mineeT) %>%
      reduce(rbind)
    dfTimer <- rbind(dfTimer, dfTimerC)
    t <-  t+1
  }
  save(dfTimer, file = "cache/timer.rda")
} else {
  load("cache/timer.rda")
}

#timerMelt <- melt(dfTimer, id="n")
ggplot(dfTimer, aes(x = n, y = t, color=name)) + 
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "Sample Size", y = "Time [s]", color = "Index family")
```

@@

```{r bbhGuided3, results="hide", fig.cap=paste("Final views identified in the dataset considering the seven dimensional parameter space (alpha, theta\\_jn, chi\\_tot, chi\\_p, ra, dec, distance), differing only by randomly selected starting plane."), dev = "png", dpi=300}
bbhM3 <- rescale(select(bbhD, alpha, theta_jn, chi_tot, chi_p, ra, dec, distance))
getResWithSeed <- function(s){
  set.seed(s)
  startM <- orthonormalise(matrix(runif(14),ncol = 2))
  tourRes <- save_history(bbhM3, guided_tour(splineIndex()))
}

if(!file.exists("cache/bbhGuidedSpline3.rda")){
  set.seed(2018)
  seedVals <- sample(1:10000, 6)
  bbhRes1 <- getResWithSeed(seedVals[1])
  bbhRes2 <- getResWithSeed(seedVals[2])
  bbhRes3 <- getResWithSeed(seedVals[3])
  bbhRes4 <- getResWithSeed(seedVals[4])
  bbhRes5 <- getResWithSeed(seedVals[5])
  bbhRes6 <- getResWithSeed(seedVals[6])
  save(bbhRes1, bbhRes2, bbhRes3, bbhRes4, bbhRes5, bbhRes6, file = "cache/bbhGuidedSpline3.rda")
} else {
  load("cache/bbhGuidedSpline3.rda")
}
resList <- list(as.list(bbhRes1), as.list(bbhRes2), as.list(bbhRes3),
                as.list(bbhRes4), as.list(bbhRes5), as.list(bbhRes6))
pList <- list()
idxV <- NULL
i <- 1
xmins <- c(-0.5,-0.5,-0.4,-0.9,-1,0.25)
xmaxs <- c(0.75,0.75,0.8,0.4,0.25,1.5)
ymins <- c(-0.8,-0.8,-0.7,-0.8,-1.2,-0.6)
ymaxs <- c(0.7,0.6,0.7,0.7,0.,0.9)
xscales <- c(0.25,0.25,0.24,0.26,0.25,0.25)
yscales <- c(0.3,0.28,0.28,0.3,0.24,0.3)
xcents <- c(0.25,-0.25,-0.1,0.1,-0.75,1.25)
ycents <- c(0.4,-0.4,-0.4,-0.4,-0.75,0.5)
titleL <- c("a: ","b: ","c: ","d: ","f: ","e: ")
for (r in resList){
  iLast <- length(r)
  fProj <- r[[iLast]]
  dProj <- as.tibble(bbhM3 %*% fProj)
  colnames(fProj) <- c("PP1", "PP2")
  colnames(dProj) <- c("PP1", "PP2")
  fProj <- as_tibble(fProj) %>%
    add_column(label=colnames(bbhM3))
  idxV <- splines2d(dProj$PP1, dProj$PP2)
  pList[[i]] <- ggplot(dProj, aes(PP1,PP2)) + geom_point(alpha = 0.02) + theme(aspect.ratio=1)+ 
  geom_path(data=getCircle(xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x, y=y), color="grey") + 
  xlim(xmins[i], xmaxs[i]) + ylim(ymins[i], ymaxs[i]) +
  geom_segment(data=getAxes(fProj, xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x1, xend=x2, y=y1, yend=y2)) +
  geom_text(data=getAxes(fProj, xcents[i],ycents[i],xscales[i],yscales[i]), aes(x=x2, y=y2, label=label), size=2.5) +
    ggtitle(paste0(titleL[i], toString(format(idxV, digits=2))))
  i <- i+1
}
grid.arrange(pList[[1]], pList[[2]], pList[[3]],pList[[4]], pList[[6]], pList[[5]], ncol=3)
```
